Online Verification of Commutativity

### 2020-09-10 acmPaper/paper.tex
Under the page limit

\section{Introduction}
\label{introduction}
Many systems use diagrams: graphs where nodes are domains and edges are transformation functions.
A type system with coercions, for example, corresponds to a graph whose nodes are types and whose edges are coercions.
Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types~\cite{unitsOfMeasure}.
In such a system, an important correctness criterion is that the diagram \emph{commutes:} when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
With our coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
Specifically, given a variable \code{x} of type \code{meters}, applying the cast \code{(wugs) x} can be done in two ways: either \code{(wugs)(feet) x} or \code{(wugs)(miles) x}.
Which path is taken depends on the compiler; we would like the choice of paths to be semantically equivalent so the compiler is free to make a choice.

@@

This paper is about efficiently checking diagrams that arise in real systems for commutativity.
We assume a simple equivalence checker for individual transformation functions: in our type system example, for instance, it is possible to check transformation equivalence by comparing the conversion factors.
Our aim is to analyze the graph of transformations and minimize the number of times we need to perform an equivalence check.
Since diagrams may change over time in real systems, as new conversions are added, and verifying the entire system from scratch may be computationally expensive, we want an online method that only checks the impact of new edges.
In Figure~\ref{fig:typeExample}, for example, a run-time system can catch the point where the programmer adds a bad conversion definition by verifying each new conversion edge as it is created.

@@

The formal algorithms in the rest of the paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.

@@

\subsection{Na\"{i}ve Baseline Algorithm}
Our first goal is to develop a baseline (exponential) algorithm that can reason about cycles without producing an infinite set of paths.  This algorithm will first pare the structure of the graph down to remove cycles, extract the pairs of paths in the graph, and finally reason about each pair to check commutativity.  This algorithm produces the two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.

@@

The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths that omit one or more of the three phases are also accepted).

@@

Murota's main result~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
The paper describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that needs to be checked.

@@

For a set of bilinkings $R$, the closure function $cl$ is defined as:
$cl(R) = \{ s\in R_0| s$ is linearly {dependent on $F(R) \}$.
%
The closure function on $R$ basically captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
%
Using these two functions, we finally define the function $\sigma$ on a set of bilinkings $R$ as
$\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $R$.
%
We use $\sigma$ to iteratively check if a given pair is redundant.
We eliminate Bilinkings until a minimum ``spanning'' subset is reached.

>>>>

\section{Introduction}
\label{introduction}
Many systems use diagrams: graphs where nodes are domains and edges are transformation functions.
A type system with coercions, for example, corresponds to a graph whose nodes are types and whose edges are coercions.
Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types~\cite{unitsOfMeasure}.
In such a system, an important correctness criterion is that the diagram \emph{commutes}: when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
With our coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
Specifically, given a variable \code{x} of type \code{meters}, applying the cast \code{(wugs) x} can be done in two ways: either \code{(wugs)(feet) x} or \code{(wugs)(miles) x}.
Which path is taken depends on the compiler; we would like the choice of paths to be semantically equivalent so the compiler is free to make a choice.

@@

This paper is about efficiently checking commutativity in diagrams that arise in real systems.
We assume a simple equivalence checker for individual transformation functions: in our type system example, for instance, it is possible to check transformation equivalence by comparing the conversion factors.
Our aim is to analyze the graph of transformations and minimize the number of times we need to perform an equivalence check.
Since diagrams may change over time in real systems, as new conversions are added, and verifying the entire system from scratch may be computationally expensive, we want an online method that only checks the impact of new edges.
In Figure~\ref{fig:typeExample}, for example, a run-time system can catch the point where the programmer adds a bad conversion definition by verifying each new conversion edge as it is created.

@@

The algorithms in this paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.

@@

\subsection{Na\"{i}ve Baseline Algorithm}
Our first goal is to develop a baseline (exponential) algorithm that can reason about cycles without producing an infinite set of paths.  This algorithm will first pare the structure of the graph down to remove cycles, extract the pairs of paths in the graph, and finally reason about each pair to check commutativity.  This results in two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.

@@

The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths with zero or one flip are also accepted).

@@

Murota's main result~\cite{commutative} solves the batch version of \textsc{verification set}: given an acyclic diagram, it returns the minimal set of equality checks that succeed if and only if the diagram commutes.
Murota describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that need be checked.

@@

For a set of bilinkings $R$, the closure function $cl$ is defined as:
$cl(R) = \{ s\in R_0| s$ is linearly dependent on $F(R) \}$.
%
The closure function on $R$ captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
%
Using these two functions, we define the function $\sigma$ on a set of bilinkings $R$ as
$\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $R$.
%
We use $\sigma$ to iteratively check if a given pair is redundant.
We eliminate Bilinkings until we reach a minimum ``spanning'' subset.
